---
title: "Movie Recommedation System"
author: "Uma Srinivas Majji"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### What is a Recommendation System?

A recommendation system provides suggestions to the users through a filtering process that is based on user preferences and browsing history. The information about the user is taken as an input. The information is taken from the input that is in the form of browsing data. This information reflects the prior usage of the product as well as the assigned ratings. A recommendation system is a platform that provides its users with various contents based on their preferences and likings. A recommendation system takes the information about the user as an input. The recommendation system is an implementation of the machine learning algorithms.

### Loading Packages

```{r message = FALSE, warning=FALSE}
library(recommenderlab)
library(ggplot2)
library(data.table)
library(reshape2)
```

### Importing the data

```{r message = FALSE, warning=FALSE}
setwd("E:/R/Projects/Movie Recommendation System")
movie_data <- read.csv("./IMDB-dataset/movies.csv")
rating_data <- read.csv("./IMDB-dataset/ratings.csv")

head(movie_data)
summary(movie_data)

head(rating_data)
summary(rating_data)
```

### Data pre-processing

```{r message = FALSE, warning=FALSE}
movie_genre <- as.data.frame(movie_data$genres, stringsAsFactors = FALSE)
movie_genre2 <- as.data.frame(tstrsplit(movie_genre[,1], "[|]", type.convert = TRUE), 
                              stringsAsFactors = FALSE)
colnames(movie_genre2) <- c(1:10)

list_genre <- c("Action", "Adventure", "Animation", "Children", 
                "Comedy", "Crime","Documentary", "Drama", "Fantasy",
                "Film-Noir", "Horror", "Musical", "Mystery","Romance",
                "Sci-Fi", "Thriller", "War", "Western")

genre_mat1 <- matrix(0, 10330, 18)
genre_mat1[1,] <- list_genre
colnames(genre_mat1) <- list_genre

for(index in 1:nrow(movie_genre2)){
    for(col in 1:ncol(movie_genre2)){
        gen_col <- which(genre_mat1[1,] == movie_genre2[index, col])
        genre_mat1[index+1,gen_col] <- 1
    }
}
head(genre_mat1)

# remove first row
genre_mat2 <- as.data.frame(genre_mat1[-1,], stringsAsFactors = FALSE)
head(genre_mat2)

# convert from characters to integers
for(col in 1:ncol(genre_mat2)){
    genre_mat2[,col] <- as.integer(genre_mat2[,col])
}

str(genre_mat2)
```

Create a 'search matrix' that will allow us to perform an easy search of the films by specifying the genre present in our list.

```{r message = FALSE, warning=FALSE}
searchMatrix <- cbind(movie_data[,1:2], genre_mat2)
head(searchMatrix)
```

For our movie recommendation system to make sense of our ratings through recommenderlabs, we have to convert our matrix into a sparse matrix one. This new matrix is of the class ‘realRatingMatrix’. This is performed as follows:

```{r message = FALSE, warning=FALSE}
rating_matrix <- dcast(rating_data, userId~movieId, value.var = "rating", na.rm = F)
View(rating_matrix)

# removing user id's
rating_matrix <- as.matrix(rating_matrix[,-1])
View(rating_matrix)

# converting rating matrix into a recommenderlab sparse matrix
rating_matrix <- as(rating_matrix, "realRatingMatrix")
rating_matrix
```

Overview of some of the important parameters that provide us various options for building recommendation systems for movies.

```{r message = FALSE, warning=FALSE}
recommendation_model <- recommenderRegistry$get_entries(dataType = "realRatingMatrix")
names(recommendation_model)

lapply(recommendation_model, "[[", "description")
```

Implementing a single model in our R project - Item Based Collaborative Filtering

```{r message = FALSE, warning=FALSE}
recommendation_model$IBCF_realRatingMatrix$parameters
```


### Exploring similar data

Collaborative Filtering involves suggesting movies to the users that are based on collecting preferences from many other users. For example, if a user A likes to watch action films and so does user B, then the movies that the user B will watch in the future will be recommended to A and vice-versa. Therefore, recommending movies is dependent on creating a relationship of similarity between the two users. With the help of recommenderlab, we can compute similarities using various operators like cosine, pearson as well as jaccard.

```{r message = FALSE, warning=FALSE}
similarity_matrix <- similarity(rating_matrix[1:4,], method = "cosine", which = "users")
as.matrix(similarity_matrix)

image(as.matrix(similarity_matrix), main = "User's Similarities")
```


Similarity that is shared between the films.

```{r message = FALSE, warning=FALSE}
movie_similarity <- similarity(rating_matrix[, 1:4], method = "cosine", which = "items")
as.matrix(movie_similarity)

image(as.matrix(movie_similarity), main = "Movies similarity")
```

Extracting the most unique ratings.

```{r message = FALSE, warning=FALSE}
rating_values <- as.vector(rating_matrix@data)
#rating_values
unique(rating_values)
```

table of ratings that will display the most unique ratings.

```{r message = FALSE, warning=FALSE}
table(rating_values)
```



### Most viewed movies - Visualization

```{r message = FALSE, warning=FALSE}
#count views for each movie
movie_views <- colCounts(rating_matrix)

# create data frame of views
table_views <- data.frame(movie = names(movie_views), views = movie_views)

# sort by no. of views
table_views <- table_views[order(table_views$views, decreasing = TRUE), ]
table_views$title <- NA

for(index in 1:10325){
    table_views[index, 3] = subset(movie_data, 
                                   movie_data$movieId == table_views[index,1])$title
}
table_views[1:6,]
```

A bar plot for the total number of views of the top films.

```{r message = FALSE, warning=FALSE}
ggplot(table_views[1:6, ], aes(x = title, y = views)) +
    geom_bar(stat="identity", fill = 'steelblue') +
    geom_text(aes(label=views), vjust=-0.3, size=3.5) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    ggtitle("Total Views of the Top Films")
```


Heatmap of Movie Ratings

```{r message = FALSE, warning=FALSE}
image(rating_matrix[1:20, 1:25], axes = FALSE, 
      main = "Heatmap of the first 20 rows and 25 columns")
```


### Performing Data Preparation

We will conduct data preparation in the following three steps –

- Selecting useful data.
- Normalizing data.
- Binarizing the data.

We have set the threshold for the minimum number of users who have rated a film as 50 and minumum of 50 views per film.

```{r message = FALSE, warning=FALSE}
movie_ratings <- rating_matrix[rowCounts(rating_matrix) > 50,
                              colCounts(rating_matrix) > 50]
movie_ratings     # 420 users and 447 films
```

We can now delineate our matrix of relevant users as follows.

```{r message = FALSE, warning=FALSE}
minimum_movies <- quantile(rowCounts(movie_ratings), 0.98)
minimum_users <- quantile(colCounts(movie_ratings), 0.98)
image(movie_ratings[rowCounts(movie_ratings) > minimum_movies,
                    colCounts(movie_ratings) > minimum_users],
      main = "Heatmap of the top users and movies")
```


Visualization of the distribution of the average ratings per user

```{r message = FALSE, warning=FALSE}
average_ratings <- rowMeans(movie_ratings)
qplot(average_ratings, fill = I("steelblue"), col = I("red")) +
    ggtitle("Distribution of the average rating per user")
```

### Data Normalization

In the case of some users, there can be high ratings or low ratings provided to all of the watched films. This will act as a bias while implementing our model. In order to remove this, we normalize our data. Normalization is a data preparation procedure to standardize the numerical values in a column to a common scale value. This is done in such a way that there is no distortion in the range of values. Normalization transforms the average value of our ratings column to 0. We then plot a heatmap that delineates our normalized ratings.

```{r message = FALSE, warning=FALSE}
normalized_ratings <- normalize(movie_ratings)
sum(rowMeans(normalized_ratings) > 0.00001)

image(normalized_ratings[rowCounts(normalized_ratings) > minimum_movies,
                         colCounts(normalized_ratings) > minimum_users],
      main = "Normalized Ratings of the Top Users")
```


### Performing Data Binarization

Binarizing the data means that we have two discrete values 1 and 0, which will allow our recommendation systems to work more efficiently. We will define a matrix that will consist of 1 if the rating is above 3 and otherwise it will be 0.

```{r message = FALSE, warning=FALSE}
binary_minimum_movies <- quantile(rowCounts(movie_ratings), 0.95)
binary_minimum_users <- quantile(colCounts(movie_ratings), 0.95)

good_rated_films <- binarize(movie_ratings, minRating = 3)
image(good_rated_films[rowCounts(movie_ratings) > binary_minimum_movies,
                       colCounts(movie_ratings) > binary_minimum_users],
      main = "Heatmap of the top users and movies")
```

### Collabrative Filtering System

We will develop our very own Item Based Collaborative Filtering System. This type of collaborative filtering finds similarity in the items based on the people’s ratings of them. The algorithm first builds a similar-items table of the customers who have purchased them into a combination of similar items. This is then fed into the recommendation system.

Splitting the dataset into 80% training set and 20% test set

```{r message = FALSE, warning=FALSE}
sampled_data<- sample(x = c(TRUE, FALSE),
                      size = nrow(movie_ratings),
                      replace = TRUE,
                      prob = c(0.8, 0.2))
training_data <- movie_ratings[sampled_data, ]
testing_data <- movie_ratings[!sampled_data, ]
```

Building the Recommendation System using R

```{r message = FALSE, warning=FALSE}
recommendation_system <- recommenderRegistry$get_entries(dataType ="realRatingMatrix")
recommendation_system$IBCF_realRatingMatrix$parameters


recommen_model <- Recommender(data = training_data,
                              method = "IBCF",
                              parameter = list(k = 30))
recommen_model
class(recommen_model)
```

Using the getModel() function, we will retrieve the recommen_model. We will then find the class and dimensions of our similarity matrix that is contained within model_info.

```{r message = FALSE, warning=FALSE}
model_info <- getModel(recommen_model)
class(model_info$sim)
dim(model_info$sim)
top_items <- 20
image(model_info$sim[1:top_items, 1:top_items],
      main = "Heatmap of the first rows and columns")
```

We will carry out the sum of rows and columns with the similarity of the objects above 0. We will visualize the sum of columns through a distribution as follows. 

```{r message = FALSE, warning=FALSE}
sum_rows <- rowSums(model_info$sim > 0)
table(sum_rows)

sum_cols <- colSums(model_info$sim > 0)
qplot(sum_cols, fill=I("steelblue"), col=I("red")) + 
    ggtitle("Distribution of the column count")
```

We will create a top_recommendations variable which will be initialized to 10, specifying the number of films to each user. We will then use the predict() function that will identify similar items and will rank them appropriately. Here, each rating is used as a weight. Each weight is multiplied with related similarities.

```{r message = FALSE, warning=FALSE}
top_recommendations <- 10 # the number of items to recommend to each user
predicted_recommendations <- predict(object = recommen_model,
                                     newdata = testing_data,
                                     n = top_recommendations)
predicted_recommendations
```

```{r message = FALSE, warning=FALSE}
user1 <- predicted_recommendations@items[[1]] # recommendation for the first user
movies_user1 <- predicted_recommendations@itemLabels[user1]
movies_user2 <- movies_user1
for (index in 1:10){
    movies_user2[index] <- as.character(subset(movie_data,
                                               movie_data$movieId == movies_user1[index])$title)
}
movies_user2
```

```{r message = FALSE, warning=FALSE}
recommendation_matrix <- sapply(predicted_recommendations@items,
                                function(x){ as.integer(colnames(movie_ratings)[x]) }) # matrix with the recommendations for each user
#dim(recc_matrix)
recommendation_matrix[,1:4]
```
















































